Squeezing in Bose-Einstein condensates with large numbers of atoms

### 2013-10-03 large_number_squeezing_with_atoms.tex
Address comments of referee 1.
Also add the referee reports as text files plus inline comments
to the first report.

%Uncomment for PACS numbers title message
%\pacs{00.00, 20.00, 42.10}
% Keywords required only for MST, PB, PMB, PM, JOA, JOB? 
%\vspace{2pc}
%\noindent{\it Keywords}: Article preparation, IOP journals
% Uncomment for Submitted to journal title message
%\submitto{\JPA}
% Comment out if separate title page not required
\maketitle

@@

Quantum degenerate ultracold atomic sources have been used to study fundamental processes such as quantum phase transitions \cite{Leggett2001,Buluta2009}, quantum non-equilibrium thermodynamics \cite{Kinoshita:2006,Barnett:2011} and entanglement of massive particles \cite{Kheruntsyan:2005}.  Their ability to be highly isolated from their environment and controlled precisely by optical, magnetic and rf fields makes them an ideal platform for engineering and manipulating non-trivial quantum states. In quantum optics, the most common way to produce entangled states has been through squeezing, and it has been recognised that the intrinsic nonlinearities of atoms naturally lead to quadrature squeezing \cite{Kheruntsyan:2005, johnssonET2007, haineET2009, doeringET2010}.  Number squeezing can be achieved with spin squeezing techniques \cite{kitagawaET1993, gross2010, sinatra2012}, molecular dissociation of Bose-Einstein Condensates (BECs) \cite{kheruntsyanET2002,davisET2008,savage2007,ogren2010}, four-wave mixing \cite{perrin2008, ogren2009, jascula2010}, double-well potentials \cite{esteve2008, maussang2010}, or mixing quadrature-squeezed states using Ramsey interferometry \cite{haineET2009}. The creation of these non-classical atomic states has relevance for fundamental tests of quantum mechanics such as the EPR paradox \cite{ogren2010, pu2000, sorensen2001, haine2005,zhao2007}, as well as atom interferometry \cite{doeringET2010, gross2010, sorensenET1999, kuzmichET2000, meyerET2001, liebfriedET2004, roosET2006}.
While these techniques have resulted in significant experimentally detectable number squeezing, they have been limited to small numbers of atoms, typically a few hundred \cite{gross2010, esteve2008}.

@@

When considering squeezing, the Hamiltonian (\ref{eqTwoModeHamiltonian}) has typically been handled within the framework of spin squeezing, which has a long history \cite{kitagawaET1993, esteve2008,sorensenET1999, liebfriedET2004, wineland1994, steel+collett}. This approach involves exploiting the equivalence between the algebra of two harmonic oscillators and that of angular momentum, and writing
\begin{eqnarray}
\hat{J}_+ &=& \hat{a}^{\dagger} \hat{b}, \\
\hat{J}_- &=& \hat{a} \hat{b}^{\dagger}, \\
\hat{J}_x &=& \frac{1}{2} \left( \hat{J}_+ + \hat{J}_-\right), \\
\hat{J}_y &=& \frac{1}{2i} \left( \hat{J}_+ - \hat{J}_-\right), \\
\hat{J}_z &=& \frac{1}{2} \left( \hat{a}^{\dagger} \hat{a} - \hat{b}^{\dagger} \hat{b} \right).
\end{eqnarray}
Using these variables Eq.~(\ref{eqTwoModeHamiltonian}) becomes
\begin{equation}
\hat{H} = \hbar\Delta \omega(\hat{N}) \, \hat{J}_z + \frac{1}{2}\hbar\chi_+ \hat{J}_z^2 + 2\hbar\Omega \hat{J}_x - \frac{1}{4}\left(\chi_{aa} + \chi_{bb}\right)\hat{N} + \frac{1}{8} \left(\chi_{aa} + \chi_{bb} + 2\chi_{ab}\right) \hat{N}^2,
\label{eqNonlinearJosephson}
\end{equation}
where $\hat{N}=\hat{a}^\dagger\hat{a} + \hat{b}^\dagger\hat{b}$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_{aa} - \chi_{bb}\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$. For initial states which are total number Fock states, the last two terms are physically meaningless phase shifts and the first term can be removed by an appropriate choice of detuning.  In this case Eq.~\eqref{eqNonlinearJosephson} reduces to the Josephson Hamiltonian with zero detuning \cite{steel+collett}
\begin{equation}
  \hat{H}_J = \frac{1}{2} \hbar\chi_+ \hat{J}_z^2 + 2\hbar \Omega \hat{J}_x. \label{eqJosephson}
\end{equation}
The system described by $\hat{H}_J$ is then solved using angular momentum algebra.

The reduction of Eq.~\eqref{eqTwoModeHamiltonian} to the Josephson Hamiltonian has led to the common misconception that no squeezing can be generated in this system if $\chi_+ = 0$ as the effects of the $\Delta\omega$ term are usually dismissed as an irrelevant detuning.  This is not true in general for initial states other than a total number Fock state as the $\Delta\omega(\hat{N})\hat{J}_z$ term leads to a number-dependent detuning.  We demonstrate in this paper that for the case of an initial coherent state this term leads to squeezing.  A spin-echo pulse applied half-way through the experiment which exchanges the population of modes $\ket{a}$ and $\ket{b}$ (which is the protocol used in most spin squeezing experiments) also cancels the effects of this term, and in this case squeezing is only generated if $\chi_+ \neq 0$.  In this work we directly solve Eq.~\eqref{eqTwoModeHamiltonian} demonstrating that squeezing is generated (in the absence of a spin-echo pulse) even if $\chi_+=0$.
% We discuss the effects of a spin-echo pulse in Section~\ref{sec:pipulse}

@@

We solve this system analytically, and derive expressions for absolute and relative number squeezing. We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation}. If we define
\begin{eqnarray}
s &=& \sin(\theta), \label{eqsDef} \\
c &=& \cos(\theta), \\
\theta &=& \Omega(t_3 - t_2), \\
\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}}, \\
A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )], \\
A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )], \\
B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )], \\
B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )], \\
D &=& AB^* - A^*B, \label{eqDDef}
\end{eqnarray}
then the expectation values for number, number variance, and number difference variance after applying the final coupling field are given by
\begin{eqnarray}
N_a(t_3) &=& n_a c^2 + n_b s^2 + i c s D, \\
%
N_b(t_3) &=& n_a s^2 + n_b c^2 - i c s D, \\
%
{\mathrm{Var}} [ N_a](t_3) &=& n_a c^2 + n_b s^2 + 2 n_a n_b c^2 s^2 \nonumber\\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2) \nonumber \\
       && + i c s D (1-2 n_a c^2 -2 n_b s^2)  \nonumber\\
       && + 2 i c^3 s n_a (e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - e^{i(\lambda_{aa} - \lambda_{ab})} A^* B ) \nonumber \\
       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ), \\
%
{\mathrm{Var}} [ N_b](t_3) &=&  n_a s^2 + n_b c^2 + 2 n_a n_b c^2 s^2 \nonumber \\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2 ) \nonumber \\
       && + i c s D (-1+2 n_a s^2 + 2 n_b c^2)  \nonumber \\
       && + 2 i c^3 s n_b (e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - e^{i(\lambda_{bb} - \lambda_{ab})} A B^* ) \nonumber \\
       && + 2 i c s^3 n_a (e^{i(\lambda_{aa} - \lambda_{ab})} A^* B - e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* ), \\
%
{\mathrm{Var}} [ N_a - N_b](t_3) &=& n_a + n_b + 4 i c s D (n_a - n_b)(s^2 - c^2) \nonumber \\
       && + 4 c^2 s^2 (2 n_a n_b +D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} A_2^* B_2 - e^{-i(\lambda_{aa} - \lambda_{bb})} A_2 B_2^* ) \nonumber \\
       && + 4 i c s (c^2 - s^2) \left( n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B \right. \nonumber \\
       && \,\,\,\,\, \left. + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^* \right). \label{eqNumDiffVariance}
\end{eqnarray}
In the case where all the nonlinearities are equal, such that $\chi_{aa} = \chi_{ab} = \chi_{bb}$, the number variances all reduce to those of a coherent state, i.e. ${\mathrm{Var}}[N_a]=N_a$, and no number squeezing is possible. It should also be noted that, although there are three independent nonlinearities $\chi_{ij}$ in the Hamiltonian (\ref{eqTwoModeHamiltonian}), the expressions for the number variances only depend on differences. This means that number squeezing is parameterised by only two independent quantities, for example $\chi_{aa}-\chi_{ab}$ and $\chi_{aa}-\chi_{bb}$.

@@

In Figure \ref{figTwoModeAnalyticExamples} we plot a solution for the normalised number variance in mode $|a\rangle$ as a function of $\phi$, choosing a typical set of nonlinearities. This example illustrates the basic features of the system: during the final coupling period, the normalised number variance undergoes cyclic variation with a period $\tau=\pi / \Omega$, and has a series of minima. It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.  As mentioned previously, squeezing is possible in this system even when the nonlinearity in the reduced Josephson Hamiltonian $\chi_+=\chi_{aa} + \chi_{bb} - 2\chi_{ab} = 0$.  Figure~\ref{figZeroEffectiveChi} demonstrates that significant squeezing is possible in this system if a spin-echo pulse is not used.

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=8cm]{figures/modal_squeezing_effects_1D.eps}
    \caption{Effects of mode shape on squeezing. The plot shows the normalised number variance in mode $|a \rangle$ as a function of the time the final coupling pulse is applied. We compare the analytic two-mode solution (thick red), the constant density mode (solid blue), the Thomas-Fermi mode (dashed) and the Gaussian mode (dash-dotted). The nonlinearities of the various 1D simulations have been adjusted for their mode shape so that they are equivalent to that of the 0D, two-mode mode nonlinearity. Parameters used were $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \text{s}^{-1}, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$. Maximum squeezing is achieved when the atomic density profile is the most uniform.}
    \label{figModalSqueezingEffects1D}
\end{figure}

@@

>>>>

%Uncomment for PACS numbers title message
\pacs{03.75.Mn, 03.75.Kk, 42.50.Dv, 03.75.Dg}
% Keywords required only for MST, PB, PMB, PM, JOA, JOB? 
%\vspace{2pc}
%\noindent{\it Keywords}: Article preparation, IOP journals
% Uncomment for Submitted to journal title message
%\submitto{\JPA}
% Comment out if separate title page not required
\maketitle

@@

Quantum degenerate ultracold atomic sources have been used to study fundamental processes such as quantum phase transitions \cite{Leggett2001,Buluta2009}, quantum non-equilibrium thermodynamics \cite{Kinoshita:2006,Barnett:2011} and entanglement of massive particles \cite{Kheruntsyan:2005}.  Their ability to be highly isolated from their environment and controlled precisely by optical, magnetic and rf fields makes them an ideal platform for engineering and manipulating non-trivial quantum states. In quantum optics, the most common way to produce entangled states has been through squeezing, and it has been recognised that the intrinsic nonlinearities of atoms naturally lead to quadrature squeezing \cite{Kheruntsyan:2005, johnssonET2007, wuester2008, haineET2009, doeringET2010}.  Number squeezing can be achieved with spin squeezing techniques \cite{kitagawaET1993, gross2010, sinatra2012}, molecular dissociation of Bose-Einstein Condensates (BECs) \cite{kheruntsyanET2002,davisET2008,savage2007,ogren2010}, four-wave mixing \cite{perrin2008, ogren2009, jascula2010}, double-well potentials \cite{esteve2008, maussang2010}, or mixing quadrature-squeezed states using Ramsey interferometry \cite{haineET2009}. The creation of these non-classical atomic states has relevance for fundamental tests of quantum mechanics such as the EPR paradox \cite{ogren2010, pu2000, sorensen2001, haine2005,zhao2007}, as well as atom interferometry \cite{doeringET2010, gross2010, sorensenET1999, kuzmichET2000, meyerET2001, liebfriedET2004, roosET2006}.
While these techniques have resulted in significant experimentally detectable number squeezing, they have been limited to small numbers of atoms, typically a few hundred \cite{gross2010, esteve2008}.

@@

When considering squeezing, the Hamiltonian (\ref{eqTwoModeHamiltonian}) has typically been handled within the framework of spin squeezing, which has a long history \cite{kitagawaET1993, esteve2008,sorensenET1999, liebfriedET2004, wineland1994, steel+collett}. This approach involves exploiting the equivalence between the algebra of two harmonic oscillators and that of angular momentum, and writing
\begin{eqnarray}
\hat{J}_+ &=& \hat{a}^{\dagger} \hat{b}, \\
\hat{J}_- &=& \hat{a} \hat{b}^{\dagger}, \\
\hat{J}_x &=& \frac{1}{2} \left( \hat{J}_+ + \hat{J}_-\right), \\
\hat{J}_y &=& \frac{1}{2i} \left( \hat{J}_+ - \hat{J}_-\right), \\
\hat{J}_z &=& \frac{1}{2} \left( \hat{a}^{\dagger} \hat{a} - \hat{b}^{\dagger} \hat{b} \right).
\end{eqnarray}
Using these variables Eq.~(\ref{eqTwoModeHamiltonian}) becomes
\begin{equation}
\hat{H} = \hbar\Delta \omega(\hat{N}) \, \hat{J}_z + \frac{1}{2}\hbar\chi_+ \hat{J}_z^2 + 2\hbar\Omega \hat{J}_x - \frac{1}{4}\left(\chi_{aa} + \chi_{bb}\right)\hat{N} + \frac{1}{8} \left(\chi_{aa} + \chi_{bb} + 2\chi_{ab}\right) \hat{N}^2,
\label{eqNonlinearJosephson}
\end{equation}
where $\hat{N}=\hat{a}^\dagger\hat{a} + \hat{b}^\dagger\hat{b}$ is the total number operator, $\Delta \omega(\hat{N}) = (\hat{N}-1)\left(\chi_{aa} - \chi_{bb}\right)/2$ and $\chi_+=\chi_{aa}+\chi_{bb}-2\chi_{ab}$. In this form, the last two terms in Eq.~(\ref{eqNonlinearJosephson}) are proportional to powers of the number operator, and so will commute with the number variance operators for state $|a\rangle$ and $|b\rangle$. This means they cannot cause number squeezing of any type, and can be dropped in the search for a minimal, squeezing-generating Hamiltonian. Further, in the case where the system is in a total number Fock state, the first term in Eq.~(\ref{eqNonlinearJosephson}) can be removed by an appropriate choice of detuning. In this case, dropping the first and last two terms transforms Eq.~(\ref{eqNonlinearJosephson}) into the Josephson Hamiltonian with zero detuning \cite{steel+collett}:
\begin{equation}
  \hat{H}_J = \frac{1}{2} \hbar\chi_+ \hat{J}_z^2 + 2\hbar \Omega \hat{J}_x. \label{eqJosephson}
\end{equation}
The system described by $\hat{H}_J$ is then solved using angular momentum algebra.

The reduction of Eq.~\eqref{eqTwoModeHamiltonian} to Eq.~\eqref{eqJosephson} has led to the common misconception that no squeezing can be generated in this system if $\chi_+ = 0$ as the effects of the $\Delta\omega$ term are usually dismissed as an irrelevant detuning.  This is not true in general for initial states other than a total number Fock state as the $\Delta\omega(\hat{N})\hat{J}_z$ term leads to a number-dependent detuning.  We demonstrate in this paper that for the case of an initial coherent state this term leads to squeezing.  A spin-echo pulse applied half-way through the experiment which exchanges the population of modes $\ket{a}$ and $\ket{b}$ (which is the protocol used in most spin squeezing experiments) also cancels the effects of this term, and in this case squeezing is only generated if $\chi_+ \neq 0$.  In this work we directly solve Eq.~\eqref{eqTwoModeHamiltonian} demonstrating that squeezing is generated (in the absence of a spin-echo pulse) even if $\chi_+=0$.
% We discuss the effects of a spin-echo pulse in Section~\ref{sec:pipulse}

@@

We solve this system analytically, and derive expressions for absolute and relative number squeezing. We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation}. If we define
\begin{eqnarray}
\tau &=& t_3 - t_2, \\
\theta &=& \Omega \tau, \\
s &=& \sin(\theta), \label{eqsDef} \\
c &=& \cos(\theta), \\
\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}}, \\
A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )], \\
A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )], \\
B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )], \\
B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )], \\
D &=& AB^* - A^*B, \label{eqDDef}
\end{eqnarray}
then the expectation values for number, number variance, and number difference variance after applying the final coupling field are given by
\begin{eqnarray}
N_a(t_3) &=& n_a c^2 + n_b s^2 + i c s D, \\
%
N_b(t_3) &=& n_a s^2 + n_b c^2 - i c s D, \\
%
{\mathrm{Var}} [ N_a](t_3) &=& n_a c^2 + n_b s^2 + 2 n_a n_b c^2 s^2 \nonumber\\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2) \nonumber \\
       && + i c s D (1-2 n_a c^2 -2 n_b s^2)  \nonumber\\
       && + 2 i c^3 s n_a (e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - e^{i(\lambda_{aa} - \lambda_{ab})} A^* B ) \nonumber \\
       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ), \\
%
{\mathrm{Var}} [ N_b](t_3) &=&  n_a s^2 + n_b c^2 + 2 n_a n_b c^2 s^2 \nonumber \\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2 ) \nonumber \\
       && + i c s D (-1+2 n_a s^2 + 2 n_b c^2)  \nonumber \\
       && + 2 i c^3 s n_b (e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - e^{i(\lambda_{bb} - \lambda_{ab})} A B^* ) \nonumber \\
       && + 2 i c s^3 n_a (e^{i(\lambda_{aa} - \lambda_{ab})} A^* B - e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* ), \\
%
{\mathrm{Var}} [ N_a - N_b](t_3) &=& n_a + n_b + 4 i c s D (n_a - n_b)(s^2 - c^2) \nonumber \\
       && + 4 c^2 s^2 (2 n_a n_b +D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} A_2^* B_2 - e^{-i(\lambda_{aa} - \lambda_{bb})} A_2 B_2^* ) \nonumber \\
       && + 4 i c s (c^2 - s^2) \left( n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B \right. \nonumber \\
       && \,\,\,\,\, \left. + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^* \right). \label{eqNumDiffVariance}
\end{eqnarray}
In the case where all the nonlinearities are equal, such that $\chi_{aa} = \chi_{ab} = \chi_{bb}$, the number variances all reduce to those of a coherent state, i.e. ${\mathrm{Var}}[N_a]=N_a$, and no number squeezing is possible. It should also be noted that, although there are three independent nonlinearities $\chi_{ij}$ in the Hamiltonian (\ref{eqTwoModeHamiltonian}), the expressions for the number variances only depend on differences. This means that number squeezing is parameterised by only two independent quantities, for example $\chi_{aa}-\chi_{ab}$ and $\chi_{aa}-\chi_{bb}$.

@@

In Figure \ref{figTwoModeAnalyticExamples} we plot a solution for the normalised number variance in mode $|a\rangle$ as a function of $\phi$, choosing a typical set of nonlinearities. This example illustrates the basic features of the system: during the final coupling period, the normalised number variance undergoes cyclic variation with a period $T=\pi / \Omega$, and has a series of minima. It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.  As mentioned previously, squeezing is possible in this system even when the nonlinearity in the reduced Josephson Hamiltonian $\chi_+=\chi_{aa} + \chi_{bb} - 2\chi_{ab} = 0$.  Figure~\ref{figZeroEffectiveChi} demonstrates that significant squeezing is possible in this system if a spin-echo pulse is not used.

@@

In the multimode case, the single mode definition for number squeezing, $\langle\mathrm{Var}[\hat{N}_i] \rangle / \langle \hat{N}_i \rangle$, where $\hat{N}_i$ is the number operator for state $|i\rangle \in \{|a\rangle, |b\rangle \}$, is no longer adequate. In the multimode case we define the number squeezing in states $|i\rangle$ to be
\begin{equation}
\mathrm{Squeezing} = \frac{\mathrm{Var} \left[ \sum_j \hat{N}_i(x_j) \right]} { \sum_j \hat{N}_i(x_j) }
\label{eqMultimodeSqueezingDefinition}
\end{equation}
where the sums are over spatial modes defined as delta functions in position, and there is seperate number operator $\hat{N}_i(x_j)$ for each mode $j$ in state $|i\rangle$.

@@

\begin{figure}
    \centering
    \includegraphics[width=8cm]{figures/modal_squeezing_effects_1D.eps}
    \caption{Effects of mode shape on squeezing. The plot shows the normalised number variance defined by Eq.~(\ref{eqMultimodeSqueezingDefinition}) in mode $|a \rangle$ as a function of the time the final coupling pulse is applied. We compare the analytic two-mode solution (thick red), the constant density mode (solid blue), the Thomas-Fermi mode (dashed) and the Gaussian mode (dash-dotted). The nonlinearities of the various 1D simulations have been adjusted for their mode shape so that they are equivalent to that of the 0D, two-mode mode nonlinearity. Parameters used were $n_a = n_b =1 \times 10^5$, $\chi_{aa}=0.03 \text{s}^{-1}, \chi_{ab}=\chi_{bb}=0, \tau_{\mathrm{hold}}=3\times 10^{-4}$s, $\phi=4.0$. Maximum squeezing is achieved when the atomic density profile is the most uniform.}
    \label{figModalSqueezingEffects1D}
\end{figure}

@@

\bibitem{wuester2008} W{\"{u}}ster S, Dabrowska-W{\"{u}}ster B J, Scott S M, Close J D and Savage C M 2008
 Quantum-field dynamics of expanding and contracting Bose-Einstein condensates \emph{Phys. Rev.} A {\bf 77} 023619
